---
title: "STAT 413/613 HW 3: Lists and Iteration"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: no
    toc_depth: 4
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align  = "center",
                      fig.height = 3, 
                      fig.width  = 5)
```

# Instructions {-}

1. Clone this homework repo to your homework directory as a new repo.
2. Rename the starter file under the analysis directory as `hw_01_yourname.Rmd` and use it for your solutions.   
3. Modify the "author" field in the YAML header.  
4. Stage and Commit R Markdown and HTML files (no PDF files).   
5. **Push both .Rmd and HTML files to GitHub**.   
- Make sure you have knitted to HTML prior to staging, committing, and pushing your final submission.  
6. **Commit each time you answer a part of question, e.g. 1.1**   
7. **Push to GitHub after each major question**   
8. When complete, submit a response in Canvas  
    
- Only include necessary code to answer the questions.
- Most of the functions you use should be from the tidyverse. Too much base R will result in point deductions.
- Use Pull requests and or email to ask me any questions. If you email, please ensure your most recent code is pushed to GitHub.

- Learning Outcomes:
    - Manipulate vectors in base-R syntax.
    - Apply iterations with for loops.

**Libraries**
```{r}
library(tidyverse)
library(stats)
```


# Working with Lists

Because of their generality, lists (or list-like objects) are often the output of many statistical procedures in R. The file `fpout.RDS` in the data folder contains sample output from using [fitPoly](https://cran.r-project.org/package=fitPoly), a statistical
method to quantify properties of locations on the [genome](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5915555/).  

1. Use `readRDS()` and a relative path to read this data into R.
```{r}
fpout <- readRDS(file = "../data/fpout.RDS")
```

2. Use appropriate functions to obtain the length of the list and then the vector of the names from the list.
```{r}
length(fpout)
names(fpout)
```

3. The `diploscores` element does not provide any information. Remove it from the list.
```{r}
fpout_01 <- fpout[-c(5)]

# check again
length(fpout_01)
names(fpout_01)

# second method
#fpout$diploscores <- NULL

# third method
#[[5]] <- NULL
  
```

4. The `scores` element contains the output most users would want. The variables in `scores` called `P0`, `P1`, `P2`, `P3`, `P4`, `P5`, and `P6` contain "posterior probabilities" for each individual for values    `0`, `1`, `2`, `3`, `4`, `5`, and `6` (respectively). 

- A quantity useful in Bayesian analysis is called the "posterior mean," which in this case is calculated as follows:
  + posterior_mean = `(P0 * 0) + (P1 * 1) + (P2 * 2) + (P3 * 3) + (P4 * 4) + (P5 * 5) + (P6 * 6)`.
- Calculate the posterior mean for each individual and add this as a new variable in the `scores` data frame.
```{r}
fpout_01$scores %>% 
  mutate(Posterior_Mean = (P0 * 0) + (P1 * 1) + (P2 * 2) + (P3 * 3) + (P4 * 4) + (P5 * 5) + (P6 * 6)) -> scores_01

head(scores_01)
```
5. Use a `map*()` function to identify the names of the variables in the `scores` data frame that are *not* of type double. 
```{r}
map(scores_01, typeof) 
# $marker, $MarkerName, and $SampleName, are not dbl type
```
- Create a new element called `col_means` in the list that contains just the column means of all the *double* variables in the `scores` data frame.
```{r}
scores_01 %>%
 select(ratio:Posterior_Mean) -> scores_02
col_means <- list(col_means = map_dbl(scores_02, mean, na.rm = TRUE))

#  Create a new list called `col_means` 
print(col_means)

# adding to the existing list to "fpout_02"
append(fpout_01, col_means) -> fpout_02
# test whether is successful or not
length(fpout_02)
```

6. Demonstrate three different ways to extract the `col_means` element from the list. The extracted element should *not* be a list. 
```{r}
# first way
fpout_02[[7]]

# second way
fpout_02$col_means

# third method
fpout_02[["col_means"]]


# we can always check with str()
```
- Show two ways to extract the third element of `col_means`
```{r}
# first way
fpout_02$col_means[c(3)]

#second way
fpout_02$col_means[c("P1")]

#third way (optional)
fpout_02$col_means[c(-1, -2, -4, -5, -6, -7, -8, -9, -10, -11, -12)]
```

# For Loops

Consider the recursive sequence defined by
$$
x_n = x_{n-1} + \frac{|x_{n-3} - x_{n-2}|}{4}.
$$
That is, element $n$ is the sum of element $n-1$ and the absolute value of the difference between between elements $n-3$ and $n-2$ divided by two. For example, if we let $x_1 = 3$, $x_2 = 1$, and $x_3 = 10$, then $x_4$ is
$$
x_4 = 10 + \frac{|3 - 1|}{4} = 11.
$$

1. Write a function called `calcn()` that takes as input a vector `x` containing the first three elements of this sequence and an integer `n` denoting the final element of the sequence to calculate.  
- `calcn()` should return element `n`.
- Include error checking to to ensure the inputs are of the correct length and type and n is greater than 0.

For example, in my implementation of `calcn()`, I obtained the following: (see HTML)

* Proper Function structure
```{r}
calcn <- function(x, n){
  nums <- vector(mode = "integer", length = n) # set nums type
  for(i in seq_along(nums)){
    if (i <= 3){
      nums[i] <- x[i] # 1, 2, 3 can not be calculated so keep it
    }
    else {
      nums[i] <- nums[i-1] + (abs(nums[i-3] - nums[i-2]))/4 # start working formula with 4 
    }
    
    }
    return(nums[n]) # return Xn in formula
}

calcn(x = c(2, 4, 3), n = 3)
calcn(x = c(2, 4, 3), n = 4)
calcn(x = c(2, 4, 3), n = 5)
calcn(x = c(2, 4, 3), n = 6)
calcn(x = c(2, 4, 3), n = 7)
```

* Error checks
```{r error=TRUE}
calcn <- function(x, n){
  if (length(x) == 3){
    if(n > 0 & is.integer(n)){
  nums <- vector(mode = "integer", length = n)
  for(i in seq_along(nums)){
    if(i <= 3){
      nums[i] <- x[i]
    }else{
      nums[i] <- nums[i-1] + (abs(nums[i-3] - nums[i-2]))/4
    }
    
    }
  return(nums[n])
    }else{
      stop("n must > 0 and n must be an integer")
      
    } # error check n 
  }else{
    stop("the length of x should be 3")
    
    } # error check x
} # created calcn function 


calcn(x = c(2, 4, 3), n = 7L) # correct x and n type
calcn(x = c(2, 4), n = 7) # length of x < 3
calcn(x = c(2, 4, 3, 5), n = 6) # length of x > 3
calcn(x = c(2, 4, 3), n = 0) # n <= 0
calcn(x = c(2, 4, 3), n = 5.53) # n is a float(dbl)

```

- Evaluate your function at the following inputs:
  + calcn(c(11,1,130), 1000L)  
  + calcn(c(11,1,130), 1L)  
  + calcn(c(7, 3, 20), 8L)
```{r}
calcn(c(11,1,130), 1000L)  
calcn(c(11,1,130), 1L)  
calcn(c(7, 3, 20), 8L)
```
  
# Question: Lists, For-loops, and `map_*()`

Lists are often used to save simulation output. You can then extract individual elements from the lists using for-loops.

Consider the $t$-test, used to test whether or not the mean of some observations is 0. We would use the following code to simulate data from a [Normal (0,1) distribution](https://en.wikipedia.org/wiki/Normal_distribution), and then use a $t$-test to test if the true mean is 0:
```{r}
x <- rnorm(n = 10, mean = 0, sd = 1)
tout <- t.test(x)
tout
```


1. The output of `t.test()` is a list-like object. Use one function to show how many elements are in the list along with their names and types.
```{r}
# only one function to show all
str(tout)

# how many elements
length(tout)

# names
names(tout)

# types
typeof(tout)
is_double(tout)
```

2. Write a for-loop to perform the following three operations on iteration `i`:
    1. Draw 10 random observations from a normal distribution with mean 0 and standard deviation 2.
    2. Run a $t$-test on these 10 observations.
    3. Save the output of the $t$-test as the `i`th element in a list called `tlist`.  
    
- Set the seed to 1 and run for 1000 iterations.
```{r}
set.seed(1)
tlist <- list()
x <- list()
for(i in 1:1000) {
  x[[i]] <- rnorm(10, mean = 0, sd = 2) # draw 10 random observations
  tlist[[i]] <- t.test(x[[i]]) # t-test the x for each time then save to the tlist
}

# randomly test some tlists
tlist[[1]]
tlist[[20]]
tlist[[300]]
```
   
3. Use the appropriate map function to extract the sample mean from each test (from 2) and pipe to an appropriate plot to show the sampling distribution of the sample mean.
- Hint: Make sure the data going into ggplot is a data frame (tibble)

* QQ plots are for testing against a known distribution or comparing two distributions to see if they are similar.
* Histogram is the more appropriate plot here.
```{r}
tlist %>%
map_dbl(~.$estimate) -> tlist_mean # extract the mean of each tlist
as_tibble(tlist_mean) -> tlist_mean01 # make sure the data save as data frame type

# plot to histogram
tlist_mean01 %>%
 ggplot(aes(x = value))+
  geom_histogram()+
  theme_bw()+
  xlab("Mean of the tlist")

#tlist_mean01 %>%
 #ggplot(aes(sample = value))+
  #geom_qq()+
  #theme_bw()

```

4. The $p$-value is an important quantity in statistics. Use a for-loop to extract the $p$-values from each test in part 2 and combine these $p$-values into a single vector called `pvec_f`. Show the first 6 values.

```{r}
pvec_f <- vector() # set pvec_f is a vector type

# start for loop
for(i in 1:1000) {
  pvec_f[[i]] <- tlist[[i]]$p.value # recall tlist$p.value to extract it
}

head(pvec_f)

```

5. Use the appropriate map function to extract the $p$-values from each test in part 2 and combine these $p$-values into a single vector called `pvec_m`. Show the first 6 values.
```{r}
map_dbl(tlist, ~.$p.value) -> pvec_m

head(pvec_m)
```

6. *Extra Credit* $p$-values have a nice property where, if the the null hypothesis is true (i.e. the mean of the observations is actually 0), then the $p$-values follow the uniform distribution. 

- Use the data from `pvec_m` to create a [QQ-plot](https://en.wikipedia.org/wiki/Q–Q_plot) and then interpret the plot with regard to whether the $p$-values exhibit a uniform distribution.  
- Ensure your plot has appropriate labels for the axes and a title. 
- Include an abline that is dashed and colored red.

**Interpretation:** If $p$-values is uniformly distributed, the null hypothesis should be true. In this plot, we can see these two lines are almost sticking together, which means this statistical evidence indicates tlist's $p$-values is uniformly distributed.
```{r}
pvec_m %>%
  as_tibble(pvec_m) -> dfpvec_m # make sure the data save as data frame type
  ggplot(data = dfpvec_m)+
  geom_qq(distribution = stats::qunif, mapping = aes(sample = pvec_m))+
  geom_abline(color = "red", linetype = "dashed", size = 1)+ # "qunif" gives the quantile function 
  theme_bw()+
  labs(x = "P-Values", y = "Sample")+
  ggtitle("p-values and uniform distribution")
```
