---
title: 'STAT 413/613 Homework on Web Data: APIs and Scraping'
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: no
    toc_depth: 4
urlcolor: blue
---

# Instructions {-}

- Write your solutions **in this starter file**. 
  + Modify the "author" field in the YAML header.
- Commit R Markdown and HTML files (no PDF files). **Push both .Rmd and HTML files to GitHub**.
  + Make sure you have knitted to HTML for your final submission.
- **Only include necessary code and data** to answer the questions.
- Most of the functions you use should be from the tidyverse. **Too much base R **will result in point deductions.
- Submit a response on Canvas that your assignment is complete on GitHub
- Feel free to use Pull requests and or email (attach your .Rmd) to ask me any questions.\

**0. Libraries**
```{r}
# This package is required for Accessing APIS (HTTP or HTTPS URLS from Web)
library(httr)
#This package exposes some additional functions to convert json/text to data frame
library(rlist)
#This package exposes some additional functions to convert json/text to data frame
library(jsonlite)
#This library is used to manipulate data
library(tidyverse)
# This puts the key into your computer’s key and credential manager for storage
library(keyring)
# Renviron file using the {usethis} package function
library(usethis)
```

# Using APIs

- Pick a website of your choice and use an API to download a data set. Convert elements of interest into a tibble and create a graph to answer a question of interest.
- State the question and interpret the plot

```{r}
resp <- GET("https://opendata.arcgis.com/datasets/f08294e5286141c293e9202fcd3e8b57_1.geojson")
# http_type(resp)
# This method will tell us what is the type of response fetched from GET() call to the API.
# http_error(resp) 
# This method just verifies if the response is error free for processing
```

```{r}
# Shows raw data which is not structured and readable, if `as` is not specified, content does its best to guess which output is most appropriate.
jsonRespText<-content(resp,as="text") 
#jsonRespText
```

```{r}
# Structured data in form of R vectors and lists
jsonRespParsed <- content(resp) 
jsonRespParsed$features[[2]]
resp
```

```{r}
# Convert JSON reponse which is in text format to data frame using jsonlite package
DC2019Crime <- fromJSON(jsonRespText)
glimpse(DC2019Crime$features$properties) -> aaa
as_tibble(aaa)

aaa %>%
  select(SHIFT)
```
```{r}
modJson<-jsonRespParsed$features#. Access data element of whole list and ignore other vectors

head(modJson)
```



```{r}
bestPics <- GET(url = "https://opendata.arcgis.com/datasets/f08294e5286141c293e9202fcd3e8b57_1.geojson", query = list(t = "Crime Incidents in 2019", r = "json"))
status_code(bestPics)
headers(bestPics)
bestPics$connection
```

# IMDB List of Oscar Winners

IMDB has a list of the [Oscar Best Picture Winners](https://www.imdb.com/search/title/?count=100&groups=oscar_best_picture_winners&sort=year%2Cdesc&ref_=nv_ch_osc).

Scrape the following elements, convert the data into a tibble, tidy it, and clean it to answer the questions below: 
- Number
- Title
- Year
- MPAA Rating
- Length in minutes
- Genre
- Star Rating
- Metascore Rating
- Gross Receipts

Convert the data into a tibble, tidy it, and clean it to answer the following questions:

```{r}
status_code(bestPics)
content(bestPics, as = "text")
```

1. Which two elements are missing the most from the movies?

2. Create a plot of the length of a film and its gross, color coded by rating. 
- Does MPAA rating matter?  

3. Create a plot with a single Ordinary Least Squares smoothing line with no standard errors showing for predicting stars rating based on metacritic scores. 
- Is there a meaningful relationship in terms of the $p$-value and adjusted R-Squared?

4. Use an appropriate plot to compare the gross receipts by MPAA rating.
  + Which MPAA rating has the highest median gross receipts? 
  + Which are the R-rated movies in the overall top 10 of gross receipts?

  + Extra Credit (1 pts): Use one-way analysis of variance to assess the level of evidence for whether all ratings have the same mean gross receipts. Provide your interpretation of the results.


# Extra Credit 1 Pts

- Listen to the AI Today podcast on [Machine Learning Ops](https://podcasts.apple.com/us/podcast/ai-today-podcast-artificial-intelligence-insights-experts/id1279927057?i=1000468771571) and provide your thoughts on the following questions:  

1. Does knowing about Git and GitHub help you in understanding the podcast?  

2. How do you think the ideas of ML OPs will affect your future data science projects?  

